# 4. A Final Practice Problem 

Now for a fictional scenario. An emergency two-week randomized controlled trial of the experimental drug ZMapp is conducted to treat Ebola. (The control represents the usual standard of care for patients identified with Ebola, while the treatment is the usual standard of care plus the drug.) 

Here are the (fake) data. 

```{r read data}
d <- fread("./data/ebola_rct2.csv")
head(d)
```

You are asked to analyze it. Patients' temperature and whether they are dehydrated is recorded on day 0 of the experiment, then ZMapp is administered to patients in the treatment group on day 1. Dehydration and temperature is again recorded on day 14.

1. Without using any covariates, answer this question with regression: What is the estimated effect of ZMapp (with standard error in parentheses) on whether someone was dehydrated on day 14? What is the p-value associated with this estimate?

```{r dehydration model}
zmapp_1 <- d[ , lm(dehydrated_day14 ~ treat_zmapp)]
coeftest(zmapp_1, vcov. = zmapp_1$vcovHC_)
```
**Answer: The estimated impact of the zmapp treatment is -0.23770 (0.085632) with a p-value of 0.006595. The results are statistically significant at the 0.05 level.**

2. Add covariates for dehydration on day 0 and patient temperature on day 0 to the regression from part (a) and report the ATE (with standard error). Also report the p-value.

```{r add pre-treatment measures}
zmapp_2 <- d[ , lm(dehydrated_day14 ~ treat_zmapp + dehydrated_day0 + temperature_day0)]
zmapp_2$vcovHC_ <- vcovHC(zmapp_2)
coeftest(zmapp_2, vcov. = zmapp_2$vcovHC_)
```
**Answer: The ATE is -0.165537 (0.081976) with P-value 0.046242**
3. Do you prefer the estimate of the ATE reported in part (a) or part (b)? Why? Report the results of the F-test that you used to form this opinion. 

```{r test pre-treatment variable inclusion}
zmapp_test_object <- anova(zmapp_1, zmapp_2, test='F')
zmapp_test_object
```

> **Answer: The p-value from the f-test as 6.472 × 10−7, so the means between both models are significantly different. I prefer model reported in part (b) because having those covariates gives model 2 much more explanatory power.**

4. The regression from part (2) suggests that temperature is highly predictive of dehydration. Add, temperature on day 14 as a covariate and report the ATE, the standard error, and the p-value.

```{r add pre-treatment temperature}
zmapp_3 <- d[ , lm(dehydrated_day14 ~ treat_zmapp +
dehydrated_day0 + temperature_day0 + temperature_day14)]
zmapp_3$vcovHC_ <- vcovHC(zmapp_3)
coeftest(zmapp_3, vcov. = zmapp_3$vcovHC_)[1:5, ]
```
**Answer: The ATE is -0.12010063 (0.07767979) with P-value 0.125405588**
5. Do you prefer the estimate of the ATE reported in part (b) or part (d)? What is this preference based on? 

> **Answer: I think part b is better because adding temperature_day14 is likely a bad control, and therefore it should not be included as explanatory variable in the model.**

6. Now let's switch from the outcome of dehydration to the outcome of temperature, and use the same regression covariates as in the chunk titled `add pre-treatment measures`. Test the hypothesis that ZMapp is especially likely to reduce mens' temperatures, as compared to womens', and describe how you did so. What do the results suggest?

```{r heterogeneous treatment effects}
zmapp_4 <- d[ , lm(temperature_day14 ~ treat_zmapp + temperature_day0 + dehydrated_day0 +male + (male * treat_zmapp) )]
zmapp_4$vcovHC_ <- vcovHC(zmapp_4)
stargazer(zmapp_4, type = 'text', se = list(zmapp_4$robust.se), add.lines = list( c('SE Flavor','Robust')),
column.labels = c("zmapp_4"),model.numbers=FALSE,header=F)
```
**Answer: The results suggest that treat_zmapp has a negative impact on temperature_day14. The coefficient is -0.231 (0.119). The stargazer output shows that the ATE for treatment for females is -0.231, and the ATE for males is -2.077 (p-value < 0.01). So there is a -2 degree reduction in day 14 temperature.**

7. Which group -- those that are coded as `male == 0` or `male == 1` have better health outcomes in control? What about in treatment? How does this help to contextualize whatever heterogeneous treatment effect you might have estimated? 

```{r context for hte}
d[, mean(temperature_day14), by=.(male, treat_zmapp)]
```

> **Answer: In the control group, females have better health outcomes. In the treatment group, males also have worse health outcomes. Although the health outcomes for males are worse, the treatment effect is actually larger for them. This confirms with our findings above where we see that the male has a higher baseline average temperature in general. This gives us an idea of how the treatment works in males vs females. **

8. Suppose that you had not run the regression in part (7). Instead, you speak with a colleague to learn about heterogeneous treatment effects. This colleague has access to a non-anonymized version of the same dataset and reports that they looked at heterogeneous effects of the ZMapp treatment by each of 80 different covariates to examine whether each predicted the effectiveness of ZMapp on each of 20 different indicators of health. Across these regressions your colleague ran, the treatment's interaction with gender on the outcome of temperature is the only heterogeneous treatment effect that he found to be statistically significant. They reason that this shows the importance of gender for understanding the effectiveness of the drug, because nothing else seemed to indicate why it worked. Bolstering your colleague's confidence, after looking at the data, they also returned to his medical textbooks and built a theory about why ZMapp interacts with processes only present in men to cure. Another doctor, unfamiliar with the data, hears your colleague's theory and finds it plausible. How likely do you think it is ZMapp works especially well for curing Ebola in men, and why? (This question is conceptual can be answered without performing any computation.)

> **Answer: Performing repeated tests as above sounds like fishing expedition, which could substantially increases the chances of having a falsely significant result. It is likely that one covariate will be significant at the 0.05 level just by chance. **

9. Now, imagine that your colleague's fishing expedition did not happen, but that you had tested this heterogeneous treatment effect, and only this heterogeneous treatment effect, of your own accord. Would you be more or less inclined to believe that the heterogeneous treatment effect really exists? Why?

> **Answer: I would be more inclined to believe it. With independent test, the result is more reliable, and it is more likely that the results in fact are significant. **

10. Now, imagine that your colleague's fishing expedition **did** happen, but that you on your own tested this and only this HTE, discover a positive result and conclude there is an effect. How does your colleague's behavior change the interpretation of your test? Does this seem fair or reasonable?

>  **Answer: In this case, I would not trust the results. Given that the data is the same, the test could still be significant even if it is a false positive. I would suggest to test this by collecting data in a new trial.**
